---
title: 'Homework #2'
author: 
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# this chunk sets the chunk options for the whole document
require(knitr)
opts_chunk$set(warning=FALSE, message=FALSE)
```

## Describing data using graphics

The following exercises use a data set called `diamonds`, which can be found in
the `ggplot2` package.  You can load it in with the following code.

```{r}
require(ggplot2)
data(diamonds)
```

1. How many observations (rows) are there in this data set?  What does each row
represent?

SOLUTION:


2. Consider the `cut` of these diamonds.  Where does this variable fall in the
taxonomy of data?  Please create a graphic of its distribution and describe it in words.

SOLUTION:


3. Next please address the same questions for the `price` variable.

SOLUTION:


4. What is the average `price` of diamonds that are of `Good` cut?  What about `Very Good`?

SOLUTION:


5. What would you expect to be the relationship between `cut` and `price`?  Please
create the appropriate graphic to visualize this relationship and describe its
structure.

SOLUTION:


6. What would you expect to be the relationship between `carat` and `price`?  Please
create the appropriate graphic to visualize this relationship and describe its
structure.

SOLUTION:


7.  When comparing diamonds of the same `carat`, what is the relationship between
`cut` and `price`? Please create the appropriate plot and interpret it.

SOLUTION:


8. When comparing diamonds of the same `cut`, what is the relationship between
`carat` and `price`? Please create the appropriate plot and interpret it.

SOLUTION:


* * *

## Data Visualization done right and wrong

Please identify two data visualizations that appear in a published scientific
journal (e.g. Science, Nature, New England Journal of Medicine).  The first graphic
should be one that strikes you as well done and the second as poorly constructed.
Please comment on the reason for your selections, what they are trying to display,
and what sort of graphic is being employed.

To submit these graphics, the easiest way will probably be to bring the graphic 
up on your computer and take a screenshot.  You can then link to them (.jpg, .png,
.gif are all good) using markdown syntax (for the reference guide, click the `?` at
the top of the markdown editor).  When you submit this .Rmd on moodle, you will 
also have to include both of those image files.




