Activity 2: Describing data
==========
## 1 and 2 variables
*Author*:

```{r, echo=FALSE, warning=FALSE, message=FALSE}
require(knitr)
opts_chunk$set(warning=FALSE,message=FALSE)
```

### National Geographic Video

The following questions concern the short video [7 Billion: Are you typical?](http://www.youtube.com/watch?v=4B2xOvKFFz4).

1.  What variables do we have data on?  What type of data is each?

SOLUTION:

2.  What numerical descriptors are referenced in the video?

SOLUTION:

3.  Write out pseudo-R-code that could have generated the statistic about the 
average height of the Dutch (that is, invent your own dataframe and column names,
but use R syntax).

SOLUTION:


* * *

### Exploratory Data Analysis

We'll be exploring the NC births data set using graphical and numerical summaries.
For more information on the data, visit [openintro.org](http://www.openintro.org/stat/data/?data=nc).

```{r, echo=FALSE}
require(mosaic)
download.file("http://www.openintro.org/stat/data/nc.RData", destfile = "nc.RData")
load("nc.RData")
```

4. How do the distributions of weights for female babies compare to that of male babies? 

SOLUTION:


5. What appears to be the cutoff age that separates mature moms from young moms?

SOLUTION:


6. Describe the relationship between the number of weeks and the amount of weight
gained.

SOLUTION:


7. Think of two variables in the data set that you expect to have *no relationship*
to one another.  Plot those variables.  Was your theory confirmed or not?

SOLUTION:

