Activity 4: Regression Diagnostics
==========
*Authors*:

The Association of Turkish Travel Agencies reports the number of foreign tourists 
visiting Turkey and tourist spending by year. You can load in the data with the 
following line of code.
  
```{r}
tourism <- read.csv("http://math.smith.edu/~bbaumer/mth241/tourism.csv")
```

1. Plot the relationship between number of tourists and spending and describe it
in words.

2. What are the explanatory and response variables?

3. Why might we want to fit a regression line to these data?

4. Do the data meet the conditions required for fitting a least squares line? Use
whatever plots are necessary.

* * *

Working together on laptops, load in the `RailTrail` data set in the `mosaic` package.

```{r echo = FALSE}
require(mosaicData)
data(RailTrail)
```


5. Describe the relationship between `volume` of ridership and the high temperature 
`hightemp`.

6. What are the explanatory and response variables?

7. Fit a regression line using `lm()`.  What is the equation for your line?

8. Examine the residual plots.  Do the conditions for regression appear reasonable?
