---
output:
  ioslides_presentation:
    incremental: true
---

# Midsemester survey results

## Time spent outside of class

```{r echo=FALSE}
survey <- read.csv("http://andrewpbray.github.io/data/midsemester-surveys.csv")
hist(survey$hrs, xlab = "number of hours/week", col = "steelblue", main = "")
```


## How challenging?

```{r echo=FALSE}
d1 <- density(survey$concept)
d2 <- density(survey$comp)
d3 <- density(survey$math)
plot(1, 0, xlim = c(0, 6), ylim = c(0, .6), type = "n", yaxt = "n",
     xlab = "challenge (high = difficult)", ylab = "", bty = "n")
COL <- c("tomato", "orange", "steelblue")
lines(d1$x, d1$y, col = COL[1], lwd = 2)
abline(v = mean(survey$concept), col = COL[1], lty = 2)
lines(d2$x, d2$y, col = COL[2], lwd = 2)
abline(v = mean(survey$comp), col = COL[2], lty = 2)
lines(d3$x, d3$y, col = COL[3], lwd = 2)
abline(v = mean(survey$math), col = COL[3], lty = 2)
legend("topright", c("concepts", "computing", "math"), col = COL, lwd = 2,
       lty = 1, bty = "n")
```


## One thing you like {.build}

- Integration of computing (Learning R)
- Group activities
- Slides online


## One thing you'd like to change {.build}

- Have homework due Sunday night.
- Have homework due Sunday night.
- Have homework due Sunday night.
- Have homework due Sunday night.

**All homeworks are now due Sunday 11:55 pm**

- Post slides earlier.


# Activity 7 Part II
- Revisit the twins data set from the quiz.
- Is there evidence that the relationship between IQs differs
between the social status groups (intercepts or slopes)?


## Model 1: slr

```{r echo=FALSE}
twins <- read.csv("http://andrewpbray.github.io/data/twins.csv")
colvec <- rep("tomato", dim(twins)[1])
colvec[twins$Social == "low"] <- "seagreen3"
colvec[twins$Social == "middle"] <- "steelblue"
plot(Foster ~ Biological, data = twins, col = colvec, pch = 16)
legend("bottomright", c("high", "middle", "low"), pch = 16, bty = "n",
       col = c("tomato", "seagreen3", "steelblue"))
# you could also make the plot using the mplot() function
# require(mosaic)
# mplot(twins)
# ggplot(data=twins, aes(x=Biological, y=Foster)) + geom_point()  + aes(colour=Social) + theme(legend.position="right") + labs(title="")
m1 <- lm(Foster ~ Biological, data = twins)
abline(m1)
```


## Model 1: slr

```{r}
summary(m1)
```


## Model 2: parallel lines

```{r echo=FALSE}
m2 <- lm(Foster ~ Biological + Social, data = twins)
plot(Foster ~ Biological, data = twins, col = colvec, pch = 16)
legend("bottomright", c("high", "middle", "low"), pch = 16, bty = "n",
       col = c("tomato", "seagreen3", "steelblue"))
abline(0, 1, col = "darkgray", lty = 2)
abline(m2$coef[1], m2$coef[2], col = "tomato", lwd = 2)
abline(m2$coef[1] + m2$coef[4], m2$coef[2], col = "seagreen3", lwd = 2)
abline(m2$coef[1] + m2$coef[3], m2$coef[2], col = "steelblue", lwd = 2)
```


## Model 2: parallel lines

```{r}
summary(m2)
```


## Model 3: 2 intercepts, 2 slopes

```{r echo=FALSE}
m3 <- lm(Foster ~ Biological * Social, data = twins)
plot(Foster ~ Biological, data = twins, col = colvec, pch = 16)
legend("bottomright", c("high", "middle", "low"), pch = 16, bty = "n",
       col = c("tomato", "seagreen3", "steelblue"))
abline(m1, col = "darkgray", lty = 2)
abline(m3$coef[1], m3$coef[2], col = "tomato", lwd = 2)
abline(m3$coef[1] + m3$coef[4], m3$coef[2] + m3$coef[6], col = "seagreen3", lwd = 2)
abline(m3$coef[1] + m3$coef[3], m3$coef[2] + m3$coef[5], col = "steelblue", lwd = 2)
```


## Model 3: 2 intercepts, 2 slopes

```{r}
summary(m3)
```


## Conclusions {.build}

- Although we find that in this data set the intercepts decrease with increasing
class, this could just be due to sampling variability and not indicative of an 
actual difference.  I.e. $\hat{\beta}_2$ and $\hat{\beta}_3$ are not statistically significant.

- This data set provides no strong indication that the slopes differ between the
social groups.

- This data is best described with a simple linear regression model.

\[ \widehat{Foster} = \hat{\beta}_0 + \hat{\beta}_1 Biological \]


## Some notes {.build}

- Interaction terms can be added whole-hog using `Foster ~ Biological * Social` or 
individually with ` Foster ~ Biological + Social + Biological:Social`.
- Categorical predictor (like `Social`) with $n$ levels get separated by R out
into $n - 1$ dummy variables that that 1 if that condition is met and 0 if not.
- These dummy variables are the estimated difference between the intercept of that
level and that of the reference level.
- An interaction term between a continuous predictor and a dummy variable (such 
as `Biological:Sociallow`) is the estimated difference between the slope for
that continuous variable for that level, and the slope for the reference level.
- Different reference levels can lead to differing p-values because you're making
different comparisons.


