\documentclass[10pt]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{fancyhdr,url,hyperref}
\usepackage{enumerate}

\oddsidemargin 0in  %0.5in
\topmargin     0in
\leftmargin    0in
\rightmargin   0in
\textheight    9in
\textwidth     6in %6in

\pagestyle{fancy}

\lhead{\textsc{MATH 141}}
\chead{\textsc{Practice}}
\lfoot{}
\cfoot{}
%\cfoot{\thepage}
\rfoot{}
\renewcommand{\headrulewidth}{0.2pt}
\renewcommand{\footrulewidth}{0.0pt}

\newcommand{\ans}{\vspace{0.25in}}
\newcommand{\R}{{\sf R}\xspace}
\newcommand{\cmd}[1]{\texttt{#1}}

\title{MATH 141:\\Intro to Probability and Statistics}
\author{Prof. Bray}
\date{Fall 2015}

\rhead{\textsc{September 16, 2014}}

\usepackage{Sweave}
\begin{document}
\input{handout-03B-prob-II-concordance}

\paragraph{Lots of babies}

Given that a family has four children, which of the following sequences of boys
and girls is most likely?
\begin{enumerate}[(a)]
\item \texttt{bbbb}
\vspace{7mm}
\item \texttt{bgbg}
\vspace{7mm}
\item \texttt{gggg}
\vspace{7mm}
\end{enumerate}


\paragraph{Education and Gender}

The table below shows the distribution of education level attained by US 
residents by gender based on data collected during the 2010 American Community 
Survey.
\begin{center}
\begin{tabular}{l p{7cm} c c }
&                                       & \multicolumn{2}{c}{\textit{Gender}} \\
\cline{3-4}
&                                                   & Male  & Female \\
\cline{2-4}
& Less than 9th grade                               & 0.07  & 0.13 \\
& 9th to 12th grade, no diploma                     & 0.10  & 0.09 \\
\textit{Highest}    & HS graduate (or equivalent)   & 0.30  & 0.20 \\
\textit{education}  & Some college, no degree       & 0.22  & 0.24 \\ 
\textit{attained}   & Associate's degree            & 0.06  & 0.08 \\
& Bachelor's degree                                 & 0.16  & 0.17 \\
& Graduate or professional degree                   & 0.09  & 0.09 \\
\cline{2-4} 
& Total                                             & 1.00  & 1.00
\end{tabular}
\end{center}
\begin{enumerate}[a]
\item What is the probability that a randomly chosen man has at least a 
Bachelor's degree?
\vspace{11mm}
\item What is the probability that a randomly chosen woman has at least a 
Bachelor's degree?
\vspace{11mm}
\item What is the probability that a man and a woman getting married both have at 
least a Bachelor's degree? Note any assumptions you must make to answer this 
question.
\vspace{11mm}
\item If you made an assumption in part~(c), do you think it was reasonable? If 
you didn't make an assumption, double check your earlier answer and then return 
to this part.
\end{enumerate}

\newpage

\paragraph{HIV in Swaziland}

Swaziland has the highest HIV prevalence in the world: 26\% of this country's 
population is infected with HIV. The ELISA test is one of the first and 
most accurate tests for HIV. For those who carry HIV, the ELISA test is 99\% 
accurate. For those who do not carry HIV, the test is 93\% accurate. If an 
individual from Swaziland has tested positive, what is the probability that he 
carries HIV?


\end{document}
