\documentclass[10pt]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{fancyhdr,url}
\usepackage{graphicx}

\oddsidemargin 0in  %0.5in
\topmargin     0in
\leftmargin    0in
\rightmargin   0in
\textheight    8.9in
\textwidth     6in %6in
%\headheight    0in
%\headsep       0in
%\footskip      0.5in

\pagestyle{fancy}

\lhead{\textsc{Prof. Bray}}
\chead{\textsc{MTH 340: Project Presentation}}
\rhead{\textsc{Spring 2014}}
\lfoot{}
\cfoot{\thepage}
\rfoot{}
\renewcommand{\headrulewidth}{0.2pt}
\renewcommand{\footrulewidth}{0.0pt}



\begin{document}
%\maketitle

\newcommand{\grademe}[2]{
\begin{table}[h!]
	\centering
	\begin{tabular}{|c|c|c|c|c|c|p{2.5in}|}
	\hline
	Criteria & Poor & Fair & Average & Good & Excellent & Comments \\
	\hline
	Volume 		& 1 & 2 & 3 & 4 & 5 & \\
	Presence 	& 1 & 2 & 3 & 4 & 5 & \\
	Emphasis 	& 1 & 2 & 3 & 4 & 5 & \\
	Eye Contact & 1 & 2 & 3 & 4 & 5 & \\
	Clarity 	& 1 & 2 & 3 & 4 & 5 & \\
	Mastery		& 1 & 2 & 3 & 4 & 5 & \\
	\hline
	\end{tabular}
	\caption{Group #1 - #2}
\end{table}
}

\section*{Gradesheet}

\paragraph{Example Form}

Your assistance is required in evaluating the presentations of your peers. Please the circle the number that best describes your perception of each group's presentation. Keep the following grading rubrics in mind. 

\begin{itemize}
	\item Volume: Are the speakers loud enough? Can you hear them even in the back of the room? 
	\item Presence: Do the speakers command the room? Do they draw your attention or do they just look like they want this to be over as soon as possible?
	\item Emphasis: Do the speakers use intonation to signal what words or concepts are most important? Or do they appear to be reciting a speech flatly?
	\item Eye Contact: Are the speakers engaged with the audience? Do they seem connected or do they seem to be going through the motions?	
	\item Clarity: Did the speakers convey a clear sense of what the project is about? If you had to tell someone what their project was about in three sentences could you do it? 
	\item Mastery: Do the speakers appear to have full command of the material? Were there obvious gaps in their knowledge of the subject matter? Were they able to answer questions coherently? 	
\end{itemize}

\grademe{1}{}
\grademe{2}{}
\grademe{3}{}




\end{document}
