---
title: 'Quiz: Simple Linear Regression'
output: html_document
---

## Part One: Twins and IQ
In the mid 20th century, a study was conducted that tracked down identical
twins where one child was raised in the home of their biological parents and
the other in a foster home.  In an attempt to answer the question of whether 
intelligence is result of nature or nurture, both children were given IQ tests.
You can load in the data with the following command:

```{r}
twins <- read.csv("http://andrewpbray.github.io/data/twins.csv")
```

The IQ of the fosterer child is recorded in `Foster`, the IQ of the biologically-
raised child in `Biological`.  The socieoeconomic status of the biological family
was also recorded in `Social`.

1) Create a plot of the relationship between the IQs of the two children and 
describe the relationship in words (there is no clear independent variable here,
so just put `Biological` on the x-axis).

```{r}
plot(Foster ~ Biological, data = twins)
```

*(2 points) There appears to be a positive linear relationship of moderate strength between
the IQ of the biological children and the IQ of the foster children.*


2. Fit a linear model to this data and have a look at the summary of the model.
Which one of the following is *false*?
    A) For each 10 point increase in the biological twin's IQ, we expect the 
    foster twin's IQ to increase on average by roughly points.
    B) The linear model is approximately $\hat{y} = 9.2 + 0.9 \times x$.
    C) Foster children with higher than average IQs will have biological twins
    with higher than average IQs as well.

```{r}
m1 <- lm(Foster ~ Biological, data = twins)
summary(m1)
```

*(2 points) A) is true: that's essentially the definition of the slope.  B) is true: that's
the equation of our fitted model.  C) is false.  Those biological twins will tend
to have higher than average IQs, but there will be variation.  Our model says
that it's possible to observe a twin pair with the foster IQ above average and 
the biological IQ below average.*


3. Is it reasonable to draw conclusions from this model?  Please assess the 
conditions for a valid model.

```{r}
plot(m1, 1)
```

*(3 points) We use the residual vs fitted plot to investigate whether the linear trend 
assumption seems reasonable.  Although the red line has a suggestion of a 
quadratic trend, it is slight and is susceptible to variation towards the edges
where there isn't much data.  The linear trend assumption is probably fine.*

```{r}
plot(m1, 2)
```

*We use the QQ plot to assess whether it's reasonable to assume that the errors
are normally distributed.  As the points follow the identity line quite well, 
this is a reasonable assumption.*

```{r}
plot(m1, 3)
```

*The scale-location plot is used to assess the constant variance assumption.  There
is no evidence of non-constant variance based on this plot.*

*The assumption of independent errors appears reasonable since it's difficult to
think of a scenario where twin pairs would be dependent.  One possibility that
we would want to rule out is that we don't have pairs with foster or biological
members that were raised in the same family (a foster family with two twins, for
example).*

*In sum, though we might want to consider a quadratic model for comparison, this
appears to be a valid model to draw inferences from.  You were not required to
comment on influential observations for this quiz, but if you did:*

```{r}
plot(m1, 5)
```

*None of our observations have usually high influence, so that is not a concern 
here.*


4. This study was used to weigh in on the question of whether IQ is a result of
nature (your genes) or nurture (your environment).  If IQ was purely a result of 
nurture, what slope would you expect to see in your linear model?  Phrase that 
hypothetical question in terms of a hypothesis test and interpret the p-value
in the context of this problem.

```{r}
summary(m1)
```

*(3 points) If IQ was a result purely of nuture, then we would expect a slope of zero. In 
testing terms, $H_0: \beta_1 = 0$, $H_A: \beta_1 \ne 0$.  The regression output
tests this hypothesis by default, yielding a p-value of `r summary(m1)$coef[2, 4]`.
As a result, we reject that null hypothesis and conclude that Biological IQ can 
be considered a significant predictor of Foster IQ.  This refutes the more general
hypothesis that IQ is entirely determined by your environment.*


5. At the other end of the argument, the distribution of IQ might be entirely
determined by genes.  Use a 95% confidence interval on your slope to assess
whether this claim is reasonable given your data set.

```{r}
ci <- confint(m1, "Biological")
```

*(3 points) The confidence interval for the slope is (`r ci[1]`, `r ci[2]`).  Because it
includes 1, this data is consistent with the argument that the distribution
of IQ is entirely due to genes.*


6. Now consider how the association between IQs would change if social class were
taken into account.  Create a plot that illustrates the relationship between
all three variables. Just by looking at the plot (and without fitting any models),
does it appear that the relationship between the IQs is the same for high social
class children as it is for low social class children?

```{r}
require(ggplot2)
ggplot(twins, aes(x = Biological, y = Foster, color = Social)) + geom_point()
```

*(2 points) If we imagine drawing the identity line through this plot, which represents equity
between biological and foster IQs, we see that in most twin pairs with biological
parents of low social class, the foster child outperforms the biological child. In
pairs with parents biological parents of high social class, there is closer parity.
It looks feasible that the relationship between the two is IQs is the the same 
for both classes (i.e. the slopes), but that the baseline IQ is different(i.e. the
intercepts).*

**Endnote**: The researcher that collected this data was named [Cyril Burt](http://en.wikipedia.org/wiki/Cyril_Burt), and his work was the subject of some controversy.

* * *

## Part Two: Education and Ethnicity in LA
Researchers in demography were interested in capturing the relationship between
the ethnic make up of neighborhoods and the level of education.  To gather data
they took a random sample of 100 neighborhoods and recorded the proportion of 
residents in each that identified as hispanic and the proportion of residents 
that had graduated from college.  The data can be loaded with the following command:

```{r}
LA <- read.csv("http://andrewpbray.github.io/data/LAed.csv")
```

1. Plot the relationship between ethnicity and education, keeping in mind the norm
of putting the independent variable on the x-axis.  Please describe the structure
in the plot as well as any unusual observations.

```{r}
plot(education ~ hispanic, data = LA)
```

*(2 points) There is a negative linear trend of moderate strength between the proportion of
hispanic residents in a neighborhood and the proportion of residents that have 
graduated from college.  There are several unusual observations.  There is an
observation that appears to have a zero recorded for both variables, suggesting
either a typo or the inclusion of a neighborhood without any residents.  There
is also a high outlier with around half hispanic residents but with an education
proportion around 1.  It is surprising that there are neighborhoods, sometimes 
several, that register as 0's or 1's on one of these metrics.*


2. Fit a linear model to this scatterplot.  Have you found a predictor that is 
statistically significant?  In your regression output, what do you look for as 
evidence and how do you interpret it?

```{r}
m2 <- lm(education ~ hispanic, data = LA)
summary(m2)
```

*(3 points) A predictor that is statistically significant is one that has a non-zero slope.
This is tested by default by running `summary(m2)`, where 
$H_0: \beta_1 = 0$, $H_A: \beta_1 \ne 0$.  The test yielded a p-value of 
`r summary(m2)$coef[2, 4]`, which leads to us rejecting the null hypothesis and 
concluding that the proportion of hispanic residents is a significant predictor
of the education level.*


3.  Was your inference conclusion based on a valid model?  Please use plots to 
assess the conditions for regression.

```{r}
plot(m2, 1)
```

*The assumption of linearity seems reasonable, however there is the possibility
that a quadratic model would yield a better fit.  We would also want to re-evaulate
after investigating the influential points.*

```{r}
plot(m2, 2)
```

*With the exception of three outliers in the tails, the QQ plot shows that the
residuals follow a nearly normal distribution.*

```{r}
plot(m2, 3)
```

*There is evidence of slightly increasing variance.*

```{r}
plot(m2, 5)
```

*Observation 35 is flagged as being an influential point.  We will want to 
consider if it appears to be a valid datum and if it belongs in this analysis.*

*It is difficult to assess indepedence since we do not know the method that was
used to collect this data.*

*In summary, the most apparent issue is some increasing variance in the residuals,
though it is not dramatic.  There is also some evidence of a quadratic trend and
one influential point (thought that need not be noted for this quiz).*


4. Say you work in state program that works to improve the number of under-
represented individuals who graduate from college.  You are deployed to one of
a 20 neighborhoods (not in this sample) in LA that is half hispanic.  You'd like 
to get a sense of the range in the graduation rate for such a neighborhood. Please
create the appropriate interval to capture that variation.

```{r}
pi <- predict(m2, data.frame("hispanic" = .5), interval = "prediction")
```

*We can capture most of the variation in the education level for a neighborhood
with 50% hispanic residents by constructing a 95% prediction interval:
(`r pi[2]`, `r pi[3]`).*

* * * 

#### Optional challenge Problem

[This is a more challenging problem that will stretch your ability.  Don't let
it frustrate you too much, but if you'd like, give it the old college try.  You 
can turn in paper and pencil work under my office door by Friday evening if
you're old school like that.]

Number 4 from the end of Chapter 2 exercises.



