---
title: 'Simple Linear Regression Quiz'
output: html_document
---

### Part One: Twins and IQ
In the mid 20th century, a study was conducted that tracked down identical
twins that were separated at birth: one child was raised in the home of their 
biological parents and the other in a foster home.  In an attempt to answer the 
question of whether intelligence is result of nature or nurture, both children 
were given IQ tests. You can load in the data with the following command:

```{r}
twins <- read.csv("http://andrewpbray.github.io/data/twins.csv")
```

The IQ of the foster child is recorded in `Foster`, the IQ of the 
biologically-raised child in `Biological`.  The socieoeconomic status of the 
biological family was also recorded in `Social`.

1) Create a plot of the relationship between the IQs of the two children and 
describe the relationship in words (there is no clear independent variable here,
so just put `Biological` on the x-axis).


2. Fit a linear model to this data and have a look at the summary of the model.
Which one of the following is *false*?
    A) For each 10 point increase in the biological twin's IQ, we expect the 
    foster twin's IQ to increase on average by roughly 9 points.
    B) The linear model (the estimated mean function) is approximately 
    $\hat{y} = 9.2 + 0.9 \times x$.
    C) Foster children with higher than average IQs (average of the foster 
    children) will have biological twins with higher than average IQs as well
    (average of the biological children).


3. Is it reasonable to draw conclusions from this model?  Please assess the 
conditions for a valid model.


4. This study was used to weigh in on the question of whether IQ is a result of
nature (your genes) or nurture (your environment).  If IQ was purely a result of 
nurture, what slope would you expect to see in your linear model?  Phrase that 
hypothetical question in terms of a hypothesis test and interpret the p-value
in the context of this problem.


5. At the other end of the argument, the distribution of IQ might be entirely
determined by genes.  Use a 95% confidence interval on your slope to assess
whether this claim is reasonable given your data set.


6. Now consider how the association between IQs would change if social class were
taken into account.  Create a plot that illustrates the relationship between
all three variables. Just by looking at the plot (and without fitting any models),
does it appear that the relationship between the IQs is the same for high social
class children as it is for low social class children?


**Endnote**: The researcher that collected this data was named [Cyril Burt](http://en.wikipedia.org/wiki/Cyril_Burt), and his work was the subject of some controversy.


* * *

### Part Two: Education and Ethnicity in LA
Researchers in demography were interested in capturing the relationship between
the ethnic make up of neighborhoods and the level of education.  To gather data
they took a random sample of 100 neighborhoods and recorded the proportion of 
residents in each that identified as hispanic and the proportion of residents 
that had graduated from college.  The data can be loaded with the following command:

```{r}
LA <- read.csv("http://andrewpbray.github.io/data/LAed.csv")
```

1. Plot the relationship between ethnicity and education, keeping in mind the norm
of putting the independent variable on the x-axis.  Please describe the structure
in the plot as well as any unusual observations.


2. Fit a linear model to this scatterplot.  Have you found a predictor that is 
statistically significant?  In your regression output, what do you look for as 
evidence and how do you interpret it?


3.  Was your inference conclusion based on a valid model?  Please use plots to 
assess the conditions for regression.


4. Say you work in state program that works to improve the number of 
under-represented individuals who graduate from college.  You are deployed to one of
20 neighborhoods (not in this sample) in LA that is half hispanic.  You'd like 
to get a sense of the range in the graduation rate for such a neighborhood. Please
create the appropriate interval to capture that variation.


* * * 

#### Optional challenge problem

[This is a more challenging problem that will stretch your ability.  Don't let
it frustrate you too much, but if you'd like, give it the old college try.  You 
can turn in paper and pencil work under my office door by Friday evening if
you're old school like that.]

Number 4 from the end of Chapter 2 exercises.



