---
title: "Properties of LSE"
output:
  ioslides_presentation:
    incremental: true
---

# Activity Review

## Interpretation of $\hat{\beta}$ {.build}

```{r echo=FALSE}
LA <- read.csv("http://andrewpbray.github.io/data/LA.csv")
LA <- transform(LA, logprice = log(price), logsqft = log(sqft))
m1 <- lm(logprice ~ logsqft + bed + city, data = LA)
m1$coef
```

- Each coefficient/slope/estimate represents the expected increase in the $y$
that is associated with an increase in the $x$ by one *with other variables held
constant*.
- Said another way, if we *control for* the other variables, a one unit increase
in $x_i$ is associated with a $\beta_i$ increase in the $y$, on average.
- Being able to control for other variables is the great strength of regression.


# Properties of Least Squares Estimates

## Inference on $\beta$ {.build}

CIs and H tests require knowledge of the sampling distribution of $\beta$. Once
transformed, each $\beta_i$ follows a t-distribution.

\[ T = \frac{\hat{\beta} - \beta}{SE(\hat{\beta})} \sim t_{n - p - 1} \]

```{r}
summary(m1)$coef
```


# Project time

