---
title: "Projects + Activity 9"
output:
  ioslides_presentation:
    incremental: true
---

# Projects

## Proposals recap {.build}

- Many groups are `merge()`ing data, which is great.  Make sure the rows align!
- Population, sample, and the observational unit.


## This week {.build}

- *Data cleaning*
- *Exploratory Data Analysis*: use the tools we started this class with (`mplot()`,
`ggplot2`, `lattice`) to explore the structure in your data.  Distribution of 
your variables (histograms, density plots), predictors vs response, relationships
between predictors.  Use scatterplots, boxplots, etc.
- *Work in Markdown!!!* be sure all your plots and data manipulation is completely
transparent and reproducible.


# Activity 9

##

1. Fit the model $price \sim sqft + bed + city$.
2. By the rules of thumb, are those two points high leverage?  Outliers?
(you can extract the hat values using `influence(m1)$hat`.)
3. Calculate the Cook's distance of those two observations using the equation:
$D_i = ((r_i^2)/(p + 1)) * ((h_{ii})/(1 - h_{ii}))$.
4. Generate the Cook's distance plot to double check that the values that you
calculated in 3 seem correct.
5. Now fit the more appropriate model, with $logprice$ and $logsqrt$ and construct
added variable plots.  What do you learn about the relative usefulness of $logsqft$
and $bed$ as predictors?


## High leverage?

```{r}
LA <- read.csv("http://andrewpbray.github.io/data/LA.csv")
m1 <- lm(price ~ sqft + bed + city, data = LA)
levs <- influence(m1)$hat
```


## High leverage?

```{r, echo=FALSE}
hist(levs)
abline(v = 2 * length(m1$coef) / nrow(LA), col = "red")
tail(sort(levs))
```


## High residual?

```{r}
e_hat <- m1$res
s <- sqrt(1/(nrow(LA) - length(m1$coef)) * sum(e_hat^2))
r <- e_hat/(s * sqrt(1 - levs))
head(r)
head(rstandard(m1))
```


## High residual?

```{r}
hist(r)
tail(sort(r))
```


## Influence

```{r}
cdist <- (r^2 / length(m1$coef)) * (levs/(1 - levs))
tail(sort(cdist))
```


## Influence

```{r}
plot(m1, 5)
```


## AVPs

```{r}
LA <- transform(LA, logprice = log(price), logsqft = log(sqft))
m2 <- lm(logprice ~ logsqft + bed + city, data = LA)
library(car)
avPlots(m2)
```


