---
title: Matrix Notation for Regression
output:
  ioslides_presentation:
    incremental: true
---

# Matrices

## Generating a matrix in R {.build}

```{r}
A <- matrix(c(2, -4, 5, 3, -1, -3), nrow = 2)
A
```

- The first argument is a vector of numbers.
- You also need to specify `nrow` or `ncol`.
- By default, `byrow = FALSE`, so it will spool the numbers down the columns
first.
- Can also change a dataframe into a matrix directly with `as.matrix()`.


## Matrix multiplication {.build}

```{r}
B <- matrix(c(2, 1, 0, 0, 1, 4), nrow = 3)
# A * B
C <- A %*% B
```

- The `*` operator will do element-wise multiplication.
- Use `%*%` for matrix multiplication (inner product).


## Other operations {.build}

*Transpose*: `t()`

```{r}
C
t(C)
```


## Other operations {.build}

*Inverse*: `solve()`

```{r}
solve(C)
C %*% solve(C)
```


# Projects


## Project Framework {.build}

- Proposal (25%)
- Oral Presentation (37.5%)
- Technical Report (37.5%)

Two components due next week:

1. **Final proposal** due next Wednesday.
2. **Final clean data set** due next Friday.


## Initial Proposals
In good shape:
- great diversity in topics.
- leads on data.

Needs work:
- what is the observational unit (i.e. the rows of the data frame)?
- what is the response?
- build a MLR.


## Final Proposal

- See website for full explanation.


## General Guidelines {.build}

- *Data types*: continuous response, continuous and categorical predictors.
- *Size*: > 200 rows, 5 - 20 predictors.
- *Sampled*: the data as a sample from some greater population to which we'd
like to make inferences.


## Additional models

- *Time series*: time series data is data that was collected over time, such as the 
price of a particular stock.

- *Logistic regression*: used when you have a binary response (e.g. predicting whether
an internet user will click on your ad).




