---
title: 'Activity: Vectors'
author:
output:
  html_document:
    css: workshop.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
require(knitr)
opts_chunk$set(warning=FALSE,message=FALSE)
```

1. Consider the vector `x <- c(12,15,8,11,24)`.  How do we create a vector of the
differences between adjacent elements?

```{r}
x <- c(12, 15, 8, 11, 24)
```

SOLUTION:


2. How can you subset the following vector to exclude both elements less than 10 
and all even elements? (hint: `%%`)

```{r}
x <- c(12, 15, 8, 11, 24)
```


SOLUTION:


3.  Build the following vectors using the colon operator: `:`.

\[
\{-1, 0, 1, 2\} \\
\{2, 1, 0, -1\}
\]

SOLUTION:


4.   Build the following vectors using `seq()`.

\[
\{-1, 0, 1, 2\} \\
\{2, 4, 6, 8, 10, 12\} \\
\{2, 3.5, 5, 6.5\}
\]

SOLUTION:


5.   Build the following vectors using `rep()`.

\[
\{1, 2, 3, 1, 2, 3, 1, 2, 3\} \\
\{1, 1, 2, 2, 3, 3\} \\
\{1, 1, 2, 3, 3\} \\
\{1, 2, 3, 1, 2, 3, 1\} \\
\{1, 1, 2, 2, 3, 3, 1, 1, 2, 2, 3, 3, 1, 1, 2, 2, 3, 3\}
\]

SOLUTION:

