---
title: 'Lab 1: Intro to data'
author: "Key"
output:
  html_document:
    highlight: pygments
  pdf_document: default
---

```{r setup, include=FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(oilabs)
```

* * *


## Exercises:

#### Exercise 1: 

There are 20,000 cases and 9 variables in this data set. `genhlth` is ordinal
categorical; `exerany`, `hlthplan`, and `smoke100` are categorical; `height`, 
`weight`, `wtdesire`, and `age` are continuous, though they've been measured
discretely; `gender` is categorical.


#### Exercise 2: 

```{r ex2}
cdc %>%
  group_by(exerany) %>%
  summarize(median_wt = median(weight))
```

The median weight of subjects who exercised in the last month and those who did
not were the same: 165 lb.


#### Exercise 3: 

```{r ex3}
cdc %>%
  select(gender) %>%
  table()

cdc %>%
  select(gender) %>%
  table()/20000

cdc %>%
  select(genhlth) %>%
  table()/20000
```

There are 9569 males in the sample. 23.3% of the sample reports being in excellent
health. This exercise was poorly worded, so you may have also looked at the joint 
distribution of these two variables, which is fine.


#### Exercise 4: 

```{r ex4}
cdc %>%
  select(gender, smoke100) %>%
  table()
```

`r round(5022 / (5022 + 4547))` percent of men are smokers as are `r round(4419 / (4419/6012))`
 percent of women. `r round(5022 / 20000)` percent of the sample are male smokers
 while `r round(4419 / 20000)` percent of the sample are female smokers. The first
 pair of statistics is more useful in determining if men or women are more likely
 to be smokers since it controls for the possibility that there are different
 numbers of men and women in the sample.
 

#### Exercise 5: 

```{r ex5}
qplot(x = gender, fill = smoke100, data = cdc)
qplot(x = gender, fill = smoke100, data = cdc, position = "fill")
```

The first plot shows only the counts of smokers/non-smokers in each gender while
the second plot controls for the number of males and females and shows the 
proportion of smokers/non-smokers in each group. Similar to the reasoning in 
exercise 4, this second plot gives better insight into the relative proportion
of smokers in the two groups.


#### Exercise 6:

```{r ex6}
cdc <- cdc %>%
  mutate(bmi = (weight / height^2) * 703)

qplot(y = bmi, x = exerany, data = cdc, geom = "boxplot")
```

Body mass index is likely effected by the level of physical activity so I chose
to compare `bmi` to the `exerany` variable. The distribution of BMI in both
groups is right skewed with many outliers. The exercisers had a lower median and
Q3 BMI as well as a slightly lower Q1 BMI. This agrees with the intuition that 
those that exercise more will have a lower BMI, on average. There is a notable
outlier, though: the highest BMI belongs to an exerciser.


#### Exercise 7:

```{r}
under23_and_smoke <- cdc %>%
  filter(smoke100 == TRUE & age < 23)
dim(under23_and_smoke)
```

There are 620 subjects that fit the criteria of being smokers and under the age
of 23.


* * *

## On your own:

#### 1: 

```{r oyo1}
qplot(x = weight, y = wtdesire, data = cdc, geom = "point")
```

There is a positive and roughly linear association between subjects' weight and
their desired weight. Notably, most of the points fall below the identity line,
suggesting most people desire a weight that is lower than their recorded weight.

*please provide interp*


#### 2: 

```{r oyo2}
cdc <- cdc %>%
  mutate(wdiff = wtdesire - weight)
```

#### 3: 

`wdiff` is a continuous variable that has been measured discretely. A positive
`wdiff` means that person wishes to gain weight; a negative `wdiff` means that
person wishes to lose weight; and a `wdiff` of zero indicates that person is 
happy with their current weight.


#### 4: 

```{r oyo4}
qplot(x = wdiff, data = cdc, geom = "histogram")
summary(cdc$wdiff)
```

The `wdiff` data is centered at a median of -10 lb, meaning that the majority of
people wish to lose weight. The IQR is 21 lb, though there is a low outlier at 
-300 lb and a high outlier at 500 lb. Excluding the outliers, the data appears
to have a slight left skew.


#### 5: 

```{r oyo5}
qplot(x = gender, y = wdiff, data = cdc, geom = "boxplot")
```

The outliers dominate the range of the plot, so it's difficult to compare the
two groups. Looking closely, women have a lower Q1 and median `wdiff` than do men,
suggesting more women wish to lose more weight than men.

Note: these plots can be improve by restricting the limits of the y-axis using 
the following additional argument.

```{r oyo5b}
qplot(x = gender, y = wdiff, data = cdc, geom = "boxplot", ylim = c(-100, 100))
```


#### 6:

```{r oyo6}
within_one_sd <- cdc %>%
  select(weight) %>%
  filter(weight < (mean(weight) + sd(weight)) &
           weight > (mean(weight) - sd(weight)))
dim(within_one_sd)
```

`r 14152 / 20000` percent of the sample falls within one standard deviation of 
the mean weight.

FYI: you can create a single pipeline to create that proportion using something
we haven't seen yet - the `ifelse` function.

```{r oyo6b}
cdc %>%
  mutate(within_one_sd = ifelse(weight < (mean(weight) + sd(weight)) & 
                                  weight > (mean(weight) - sd(weight)),
                                "yes", "no")) %>%
  group_by(within_one_sd) %>%
  summarize(n = n()) %>%
  mutate(prop = n / sum(n))
```

