---
title: "Simple Linear Regression II"
output:
  html_document:
    css: ../lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```


## Twins and IQ
In the mid 20th century, a study was conducted that tracked down identical
twins that were separated at birth: one child was raised in the home of their 
biological parents and the other in a foster home.  In an attempt to answer the 
question of whether intelligence is result of nature or nurture, both children 
were given IQ tests.

This data is in a `.csv` file, which is a very general format for data and the
format that your project data will be in. You can load our usual packages as
well as the data with the following commands:

```{r, message = FALSE}
library(dplyr)
library(ggplot2)
twins <- read.csv("http://andrewpbray.github.io/data/twins.csv")
```

When you're working on your project, you'll be able to read it your data into
your report by calling the same function, `read.csv()`, but using the path 
`"../data/data.csv"`.

Back to the data at hand, since this data was read in from a `.csv` file, there
is no help file associated with it, so I'll just tell you that the IQ of the 
foster child is recorded in `Foster` and the IQ of the biologically-raised child in
`Biological`.  The socieoeconomic status of the biological family was also 
recorded in `Social`.

1. Create a plot of the relationship between the IQs of the two children and
   describe the relationship in words (there is no clear independent variable 
   here, so just put `Biological` on the x-axis).


2. Fit a linear model to this data and have a look at the summary of the model.
Which one of the following is *false* and why?
    - For each 10 point increase in the biological twin's IQ, we expect the 
    foster twin's IQ to increase on average by roughly 9 points.
    - The linear model (the estimated mean function) is approximately 
    $\hat{y} = 9.2 + 0.9 \times x$.
    - Foster children with higher than average IQs (average of the foster 
    children) will have biological twins with higher than average IQs as well
    (average of the biological children).


3. Is it reasonable to draw conclusions from this model?  Please assess the 
conditions for a valid model.


4. This study was used to weigh in on the question of whether IQ is a result of
nature (your genes) or nurture (your environment).  If IQ was purely a result of 
nurture, what slope would you expect to see in your linear model?  Phrase that 
hypothetical question in terms of a hypothesis test and interpret the p-value
in the context of this problem.


5. At the other end of the argument, the distribution of IQ might be entirely
determined by genes.  Use a 95% confidence interval on your slope to assess
whether this claim is reasonable given your data set.


6. Now consider how the association between IQs would change if social class were
taken into account.  Create a plot using the code below to illustrate the 
relationship between all three variables. Just by looking at the plot (and 
without fitting any models), does it appear that the relationship between the 
IQs is the same for high social class children as it is for low social class 
children?

```{r ex6}
qplot(x = Biological, y = Foster, color = Social, data = twins, geom = "point")
```

**Endnote**: The researcher that collected this data was named [Cyril Burt](http://en.wikipedia.org/wiki/Cyril_Burt), and his work was the subject of 
some controversy.