\documentclass[11pt]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{fancyhdr,url}
\usepackage{graphicx}

\oddsidemargin 0in  %0.5in
\topmargin     0in
\leftmargin    0in
\rightmargin   0in
\textheight    8.9in
\textwidth     6in %6in
%\headheight    0in
%\headsep       0in
%\footskip      0.5in

\newtheorem{thm}{Theorem}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{obs}{Observation}
\newtheorem{lemma}{Lemma}
\newtheorem{claim}{Claim}
\newtheorem{definition}{Definition}
\newtheorem{question}{Question}
\newtheorem{answer}{Answer}
\newtheorem{problem}{Problem}
\newtheorem{solution}{Solution}
\newtheorem{conjecture}{Conjecture}

\pagestyle{fancy}

\lhead{\textsc{Prof. Baumer \& Horton}}
\chead{\textsc{MTH 241: Assessment}}
\rhead{\textsc{Spring 2013}}
\lfoot{}
\cfoot{\thepage}
\rfoot{}
\renewcommand{\headrulewidth}{0.2pt}
\renewcommand{\footrulewidth}{0.0pt}



\begin{document}
%\maketitle

\noindent
{\bf Group:}

\vspace{.2in}

\noindent
{\bf Members:} 


\subsection*{Presentation Evaluation}

	\begin{itemize}
\itemsep2in
		\item General (20 points): Is the topic original, interesting, and substantial -- or is it trite, pedantic, and trivial?  How much creativity, initiative, and ambition did the group demonstrate?  Is the basic question driving the project worth investigating, or is it obviously answerable without a data-based study?
		\item Design (15 points):  Are the variables chosen appropriately and defined clearly, and is it clear how they were measured/observed?  Can the effects of lurking variables be controlled for?  Is there sufficient data to make meaningful conclusions? 
		\item Analysis (15 points):  Are the chosen analyses appropriate for the variables/relationships under investigation, and are the assumptions underlying these analyses met?  Do the analyses involve fitting and interpreting a multiple regression model?  Are the analyses carried out correctly?  Is there an effective mix of graphical, numerical, and inferential analyses?  Did the group make appropriate conclusions from the analyses, and are these conclusions justified?
		\item Report/Technical appenix (20 points):  How effectively does the written report communicate the goals, procedures, and results of the study?  Are the claims adequately supported? How well is the report structured and organized? Does the writing style enhance what the group is trying to communicate?  How well is the report edited? Does it serve appropriately as a technical appendix? Is it readable? Does the statistical analysis justify the claims made? Clear writing, correct spelling, and good grammar are important.
		\item Oral Presentation (30 points): How effectively does the oral presentation communicate the goals, procedures, and results of the study? Do the slides help to illustrate the points being made by the speaker without distracting the audience? Do the presenters seem to be well-rehearsed? Did they properly budget their time? Does she appear to be confident in what she is saying? Are the arguments persuasive? 
	\end{itemize}



\end{document}
