---
title: "Lab 3 - Probability"
author: "Enter team name here"
output: 
  html_document: 
    highlight: pygments
---

```{r setup, include=FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

**Load Hot Hand files:**

```{r}
download.file("http://www.openintro.org/stat/data/kobe.RData", destfile = "kobe.RData")
load("kobe.RData")
```

#### Exercise 1

A streak of length one means that there was one H then a M. A streak of length
zero is a solitary M (directly following another M).

#### Exercise 2

```{r ex2}
kobe_streak <- data.frame(length = calc_streak(kobe$basket))
qplot(x = length, data = kobe_streak, geom = "histogram", binwidth = 1)
summary(kobe_streak$length)
```

The distribution of Kobe's streak is right-skewed, with a max streak of 4. His
median streak length was zero.


#### Exercise 3

```{r ex3}
coin_outcomes <- c("heads", "tails")
set.seed(302)
sim_unfair_coin <- sample(coin_outcomes, size = 100, replace = TRUE, 
                          prob = c(0.2, 0.8))
table(sim_unfair_coin)
```

Heads came up 82 times.


#### Exercise 4

```{r ex4}
shot_outcomes <- c("H", "M")
sim_basket <- sample(shot_outcomes, size = 133, replace = TRUE,
                     prob = c(0.45, 0.55))
```


* * *

## On your own:

#### 1: 

```{r oyo1}
sim_streak <- data.frame(length = calc_streak(sim_basket))
```


#### 2: 

```{r oyo2}
qplot(x = length, data = sim_streak, geom = "histogram", binwidth = 1)
summary(sim_streak$length)
```

The distribution of streak lengths for the simulated no-hot hands shooter is 
right-skewed with a typical (median) length of 0 and a maximum of 6.


#### 3: 

I would expect the distribution to be similar in shape, spread, and center (e.g.
still right-skewed), however there will be some variation since we have 
conducted the simulation afresh and each shot makes a basket probabilistically.


#### 4: 


#### 5: