---
title: "Problem Set 10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


* * *

### Part Two: Education and Ethnicity in LA
Researchers in demography were interested in capturing the relationship between
the ethnic make up of neighborhoods and the level of education.  To gather data
they took a random sample of 100 neighborhoods and recorded the proportion of 
residents in each that identified as hispanic and the proportion of residents 
that had graduated from college.  The data can be loaded with the following command:

```{r}
LA <- read.csv("http://andrewpbray.github.io/data/LAed.csv")
```

1. Plot the relationship between ethnicity and education, keeping in mind the norm
of putting the independent variable on the x-axis.  Please describe the structure
in the plot as well as any unusual observations.


2. Fit a linear model to this scatterplot.  Have you found a predictor that is 
statistically significant?  In your regression output, what do you look for as 
evidence and how do you interpret it?


3.  Was your inference conclusion based on a valid model?  Please use plots to 
assess the conditions for regression.


4. Say you work in state program that works to improve the number of 
under-represented individuals who graduate from college.  You are deployed to one of
20 neighborhoods (not in this sample) in LA that is half hispanic.  You'd like 
to get a sense of the range in the graduation rate for such a neighborhood. Please
create the appropriate interval to capture that variation.


* * * 

#### Optional challenge problem

[This is a more challenging problem that will stretch your ability.  Don't let
it frustrate you too much, but if you'd like, give it the old college try.  You 
can turn in paper and pencil work under my office door by Friday evening if
you're old school like that.]

Number 4 from the end of Chapter 2 exercises.




