---
title: "Problem Set 10"
output:
  html_document:
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### Data Analysis: Education and Ethnicity in LA

Researchers in demography were interested in capturing the relationship between
the ethnic make up of neighborhoods and the level of education.  To gather data
they took a random sample of 100 neighborhoods and recorded the proportion of 
residents in each that identified as hispanic and the proportion of residents 
that had graduated from college.  The data can be loaded with the following command:

```{r}
LA <- read.csv("http://andrewpbray.github.io/data/LAed.csv")
```

1. Plot the relationship between ethnicity and education, keeping in mind the norm
of putting the independent variable on the x-axis.  Please describe the structure
in the plot as well as any unusual observations.


2. Fit a linear model to this scatterplot.  Have you found a predictor that is 
statistically significant?  In your regression output, what do you look for as 
evidence and how do you interpret it?


3.  Was your inference conclusion based on a valid model?  Please use plots to 
assess the conditions for regression.

* * *

#### Reading

To be read next week...

* * * 

#### Practice Problems

Chapter 6, number 4 (not d), 5, and 6.




