---
title: "Problem Set 10"
output:
  html_document:
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### Data Analysis: Education and Ethnicity in LA

Researchers in demography were interested in capturing the relationship between
the ethnic make up of neighborhoods and the level of education.  To gather data
they took a random sample of 100 neighborhoods and recorded the proportion of 
residents in each that identified as hispanic and the proportion of residents 
that had graduated from college.  The data can be loaded with the following command:

```{r}
LA <- read.csv("http://andrewpbray.github.io/data/LAed.csv")
```

1. Plot the relationship between ethnicity and education, keeping in mind the norm
of putting the independent variable on the x-axis.  Please describe the structure
in the plot as well as any unusual observations.

```{r}
library(ggplot2)
qplot(x = hispanic, y = education, data = LA, geom = "point")
```

*There is a fairly linear, negative, moderately strong relationship between
proportion of hispanic residents in a neighborhood and the college graduation
rate. There are several notable outliers, including neighborhoods with graduation
rates of 1 and another with a graduation rates of zero and a hispanic propotion
also of zero.*

2. Fit a linear model to this scatterplot.  Have you found a predictor that is 
statistically significant?  In your regression output, what do you look for as 
evidence and how do you interpret it?

```{r}
m1 <- lm(education ~ hispanic, data = LA)
summary(m1)
```

*The p-value associated with the slope on hispanic is very small, so we have
evidence that it is statistically significantly different from zero. There does
appear to be a relationship between these two variables.*

3.  Was your inference conclusion based on a valid model?  Please use plots to 
assess the conditions for regression.

```{r}
qplot(x = .fitted, y = .stdresid, data = m1)
qplot(sample = .stdresid, data = m1) + geom_abline()
```

*The residuals in the residual plot are consistently centered at zero, so the 
assumption of a linear trend seems reasonable. The question of constant variance
is somewhat questionable as there is evidence of a slight increase in the variance
as the fitted values increase. In terms of if these errors and their observations
can be considered independent of one another, that is called into question by
the description of how this data was collected. These constitutes a sample of 
100 neighborhoods in LA, which may well be more than 10% of the total number of
neighborhood, which would create some dependency between the observations in this
sample. In the qqplot, the points generally do follow a line (with the exception
of some strong outliers), which indicates the assumption of normality is reasonable.*

* * *

#### Reading

To be read next week...

* * * 

#### Practice Problems

Chapter 6, number 4 (not d), 5, and 6.




